Developing a preprint culture in biology

### 2012-10-21 ms.tex
small edits

% Template for PLoS
% Version 1.0 January 2009
%
% To compile to pdf, run:
% latex ms.tex
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template

@@

Public preprints can be crucial to early-career scientists. The delay before
publication is seldom compatible with the pressure to show an impressive
publication record when applying for a scholarship or a position. Increasing the
perceived value of pre-prints as close, or equal, to journal articles will allow
young researchers to put their research outcome in the open, and build a
reputation for themselves through the diffusion of their work without fear that
this work will not be recognized by grant or job committees. In this
perspective, mixed publication systems like F1000Research should
be encouraged as they combine the best of both worlds: contributions are
initially considered preprints, but acquire full article status after positive
reviews are received.

@@

There are a number of reasons why biologists did not develop the reflex of
sharing preprints, many of which are based on common misconceptions. For
example, in contrast to other fields there is a perception in biology that
public preprints make it easier to steal ideas \cite{gin11}. However, there is
no evidence of this happening in the numerous other fields that have adopted
preprint servers, and since preprint servers create a clear record of who had
the idea first, and when, this appears to be a largely unfounded concern. In
other fields preprints serve the opposite role, they allow straightforward
establishment of precedence, letting research lay claim to an idea thus
preventing it from being ``stolen'' \cite{gin11}

Another major concern is based on a certain interpretation of the Ingelfinger
rule: scientists should not publish the same manuscript twice \cite{alt96}. A
preprint is simply a document that allows ideas to spread and be discussed, it
is not yet formally validated by the peer-review system. This is why the
majority of publishers do not see arXiv and similar services as a violation of
the Ingelfinger rule. Almost all of the major publishers in biology are
preprint-friendly, including: Nature Publishing Group, PLOS, BMC, PNAS,
Elsevier, and Springer \ref{table:policies}. This year, both the Ecological
Society of America and the Genetics Society of America changed their policies to
allow public preprints. \emph{Nature} even felt compelled to respond to the
rumour that they refused manuscripts submitted to arXiv by saying that
``\emph{Nature} never wishes to stand in the way of communication between
researchers. We seek rather to add value for authors and the community at large
in our peer review, selection and editing'' \cite{nat05}.  Still, a few journals
adopt a ``by default'' hostile attitude towards preprints, mostly due to the
lack of clear policy of the publishers, or perhaps because a preprint culture
has not developed in biology and the practice is still considered unusual. As an
example, Wiley-Blackwell, which publishes some of the leading journals in
biology, has no official policy on the matter.

@@

One of the advantages of figshare over arXiv for biologists is that is it not
limited to quantitative sciences. arXiv.org has sections on quantitative biology
but might not be appropriate for non-quantitative work. With its flexible
approach to preprints, figshare offers an alternative to arXiv for empirical
biologists. Furthermore, by allowing all types of content, figshare also
provides an archive for early results (\emph{e.g.}: figures, lab presentations).

@@

PeerJ (\href{https://peerj.com/}{https://peerj.com/}) is a new open access
publisher that combines both a preprint server and a peer reviewed journal.  It
is focused on the biological and medical sciences, which may help overcome the
perception that preprints do not have a place in biology. Like figshare this is
an advantage relative to arXiv for biologists doing non-quantitative work.  Also
like figshare, PeerJ allows commenting on posted preprints, improving the
potential for pre-publication dialog. In addition, preprints can be made private
if the authors choose, and shared only with selected colleagues. While this
reduces many of the benefits of preprints described above, it may allow some
researchers who would not otherwise post preprints to begin exploring the
possibility in a manner appropriate to their current circumstances.

@@

This manuscript was developed entirely as an open project on GitHub. GitHub is
one of several hosting services for collaborative development using the Git
version control system (VCS).  Git is a decentralized revision control system
created by Linus Torvalds and is used primarily to develop software, including
the Linux kernel. Git provides powerful features that allow numerous
contributers to work asynchronously on the same project, often in parallel
branches, all of which can be effortlessly merged and version controlled.  While
Git is created primarily for software development, where the use of version
control systems is standard \cite{aru12}, it is ideal for academic research
since it provides a way to collaborate on every step of the manuscript
development process, from data manipulation and analysis to writing and
revision. For example, during the development of this manuscript, each author
would clone the project (\emph{i.e.} make a personal copy), modify it, and then
merge their changes into a master branch. This takes the preprint process to an
new level, where the entire writing process is open from the beginning.

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=4in]{map.eps}
\end{center}
\caption { {\bf It can take several months, and even a few years, before a submitted
paper is officially published and citable.}  The average time to publication
varies greatly between journals and can be as low as 104 days (Evolution for
2011) to 213 (PLOS One in 2010).  Meanwhile, few people are aware of the
research that has been done since, typically, only close colleagues are given
access to the preprints. With public preprint servers, the science is
immediately available and can be openly discussed, analyzed, and integrated into
current research. It benefits both science and publishers. Both want the papers
to be well-known and cited, and public preprints make it possible to integrate
research even before publication, greatly improving immediacy.  }
\label{fig:map}
\end{figure}

>>>>

% Template for PLoS
% Version 1.0 January 2009
%
% To compile to pdf, run:
% latex ms
% bibtex ms
% latex ms
% latex ms
% dvipdf ms

@@

Public preprints can be crucial to early-career scientists. The delay before
publication is seldom compatible with the pressure to show an impressive
publication record when applying for a scholarship or a position. Increasing the
perceived value of pre-prints as close, or equal, to journal articles will allow
young researchers to put their research outcome in the open, and build a
reputation for themselves through the diffusion of their work without fear that
this work will not be recognized by grant or job committees.

@@

There are a number of reasons why biologists did not develop the reflex of
sharing preprints, many of which are based on common misconceptions. For
example, in contrast to other fields there is a perception in biology that
public preprints make it easier to steal ideas \cite{gin11}. In other fields
preprints serve the opposite role, they allow straightforward establishment of
precedence, letting research lay claim to an idea thus preventing it from being
``stolen'' \cite{gin11}

Another major concern is based on a certain interpretation of the Ingelfinger
rule: scientists should not publish the same manuscript twice \cite{alt96}. A
preprint is simply a document that allows ideas to spread and be discussed, it
is not yet formally validated by the peer-review system. This is why almost all
the major publishers in biology are preprint-friendly, including: Nature
Publishing Group, PLOS, BMC, PNAS, Elsevier, and Springer \ref{table:policies}.
This year, both the Ecological Society of America and the Genetics Society of
America changed their policies to allow public preprints. \emph{Nature} even
felt compelled to respond to the rumour that they refused manuscripts submitted
to arXiv by saying that ``\emph{Nature} never wishes to stand in the way of
communication between researchers. We seek rather to add value for authors and
the community at large in our peer review, selection and editing'' \cite{nat05}.
Still, a few journals adopt a ``by default'' hostile attitude towards preprints,
mostly due to the lack of clear policy of the publishers. As an example,
Wiley-Blackwell, which publishes some of the leading journals in biology, has no
official policy on the matter.

@@

Unlike arXiv, figshare is not limited to quantitative sciences. arXiv.org has
sections on quantitative biology but might not be appropriate for
non-quantitative work. With its flexible approach to preprints, figshare offers
an alternative to arXiv for empirical biologists. Furthermore, by allowing all
types of content, figshare also provides an archive for early results
(\emph{e.g.}: figures, lab presentations).

@@

PeerJ (\href{https://peerj.com/}{https://peerj.com/}) is a new open access
publisher that combines both a preprint server and a peer reviewed journal.  It
is focused on the biological and medical sciences, which may help overcome the
perception that preprints do not have a place in biology. Like figshare, PeerJ
allows commenting on posted preprints, improving the potential for
pre-publication dialog. In addition, preprints can be made private if the
authors choose, and shared only with selected colleagues. While this reduces
many of the benefits of preprints described above, it may allow some researchers
who would not otherwise post preprints to begin exploring the possibility in a
manner appropriate to their current circumstances.

@@

This manuscript was developed entirely as an open project on GitHub. GitHub is
one of several hosting services for collaborative development using the Git
version control system (VCS).  Git is a decentralized revision control system
created by Linus Torvalds and is used primarily to develop software, including
the Linux kernel. Git provides powerful features that allow numerous
contributers to work asynchronously on the same project, often in parallel
branches, all of which can be effortlessly merged and version controlled.  While
Git is created primarily for software development, where the use of version
control systems is standard \cite{aru12}, it is ideal for academic research
since it provides a way to collaborate on every step of the manuscript
development process. For example, during the development of this manuscript,
each author would clone the project (\emph{i.e.} make a personal copy), modify
it, and then merge their changes into a master branch. This takes the preprint
process to an new level, where the entire writing process is open from the
beginning.

@@

\begin{figure}[!ht]
\begin{center}
\includegraphics[width=4in]{map.eps}
\end{center}
\caption { {\bf It can take several months, and even a few years, before a submitted
paper is officially published and citable.} Meanwhile, few people are aware of the
research that has been done since, typically, only close colleagues are given
access to the preprints. With public preprint servers, the science is
immediately available and can be openly discussed, analyzed, and integrated into
current research. }
\label{fig:map}
\end{figure}
